ms_results

### 2023-01-20 ms_results.Rmd
new pipeline for preparing manuscript ms results and ms figures

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(tidyverse)
library(countrycode)
library(kableExtra)
library(ggpubr)
library(tidytext)

# Set color palettes 
habitat_source_colors <- c("#1B708F", # blue = marine capture
                          "#86ADA7", # light blue = freshwater capture
                          "#F37542", # medium orange = marine aquaculture
                          "#F7AF75", # light orange = inland aquaculture
                          "grey50" # unknown
                          )
```

@@

```{r trends_summary}
# Generate summary statistics for total trade trends

# Load data summary files generated by ms_results
global_export_by_source <- read.csv("outputs/global_export_by_source.csv")
all_degree_summary <- read.csv("outputs/all_degree_summary.csv")
habitat_method_degree_summary <- read.csv("outputs/habitat_method_degree_summary.csv")

# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

@@

total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)

total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)

total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)

```{r percent_export_by_source_trends}
percent_export_by_source <- read.csv("outputs/percent_export_by_source_summary.csv")
percent_export_total <- read.csv("outputs/percent_export_total_summary.csv")

percent_export_total_dom_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_dom_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "domestic exports") %>% 
  pull(percent_export)

percent_export_total_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_export_total_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "all export sources") %>% 
  pull(percent_export)

percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine capture") %>%
  pull(percent_export)

percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_marine_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine aquaculture") %>%
  pull(percent_export)

percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland capture") %>%
  pull(percent_export)

percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
percent_inland_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland aquaculture") %>%
  pull(percent_export)

```{r trends_concentration}
# Calculate import/export concentration stats

# Load data as bilateral flows by habitat and method
bilateral_habitat_method_summary <- read.csv("outputs/bilateral_habitat_method_summary.csv") %>% 
  # Set factor levels
  mutate(habitat_method = factor(habitat_method, levels = c("marine capture", "inland capture",
                                                            "marine aquaculture", "inland aquaculture",
                                                            "unknown" )))
# Set percent threshold
cumulative_percent_threshold <- 75
# Concentration in exports
## All
export_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_n_countries <- export_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
export_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

export_concentration_habitat_method_n_countries <- export_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

export_concentration_plot <- ggplot() +
  geom_line(data = export_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = export_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of exports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

# Concentration in imports
## All
import_concentration <- bilateral_habitat_method_summary %>%
  group_by(year, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_n_countries <- import_concentration %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year) %>% 
  tally()

## By habitat and method
import_concentration_habitat_method <- bilateral_habitat_method_summary %>%
  group_by(year, habitat_method, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(year, habitat_method, desc(live_weight_t)) %>%
  mutate(cumulative_percent = 100*cumsum(live_weight_t)/sum(live_weight_t))

import_concentration_habitat_method_n_countries <- import_concentration_habitat_method %>%
  filter(cumulative_percent < cumulative_percent_threshold) %>%
  group_by(year, habitat_method) %>% 
  tally()

import_concentration_plot <- ggplot() +
  geom_line(data = import_concentration_habitat_method_n_countries %>% 
              filter(habitat_method != "unknown"), aes(x = year, y = n, color = habitat_method)) +
  scale_color_manual(values = habitat_source_colors) +
  geom_line(data = import_concentration_n_countries, aes(x = year, y = n), color = "black") +
  labs(x = "", y = paste("Number of countries comprising ", cumulative_percent_threshold, "% of imports", sep = ""), color = "Source") +
  lims(y = c(0, 30)) +
  theme_bw()

ggarrange(export_concentration_plot, import_concentration_plot, common.legend = TRUE, legend = "bottom", labels = c("a", "b"))
```

```{r degree_summary_stats}
all_degree_1996 <- all_degree_summary %>%
  filter(year == 1996) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)

all_degree_2019 <- all_degree_summary %>%
  filter(year == 2019) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)

@@

@@

```{r covid_effect}
# artis %>% 
#   filter(year %in% 2019:2020) %>% 
#   group_by(year, habitat_method) %>% 
#   summarise(live_weight_t = sum(live_weight_t)) %>% 
#   pivot_wider(names_from = year, values_from = live_weight_t, names_prefix = "year") %>% 
#   mutate(change = (year2019-year2020)/year2020) %>% 
#   arrange(desc(change))
```

@@

```{r top_traders}
# Top trade tables
top_n_countries <- 5
top_n_countries_flows <- 10
top_n_regions <- 5
top_n_regions_flows <- 10

# Top exporters
## Countries
top_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture exporters
## Countries
top_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture exporters
## Countries
top_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_exporters_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_exporters_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top importers
## Countries
top_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top capture importers
## Countries
top_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_capture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_capture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top aquaculture importers
## Countries
top_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

top_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries, order_by = live_weight_t)

## Regions
top_region_aquaculture_importers_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

top_region_aquaculture_importers_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions, order_by = live_weight_t)

# Top bilateral flows
## Countries
top_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top capture bilateral flows
## Countries
top_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_capture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_capture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "capture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Top aquaculture bilateral flows
## Countries
top_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

top_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_iso3c, importer_iso3c) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_countries_flows, order_by = live_weight_t)

## Regions
top_region_aquaculture_flows_1996 <- bilateral_habitat_method_summary %>%
  filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

top_region_aquaculture_flows_2019 <- bilateral_habitat_method_summary %>%
  filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>%
  group_by(exporter_region, importer_region) %>%
  summarise(live_weight_t = sum(live_weight_t)) %>%
  ungroup() %>%
  arrange(desc(live_weight_t)) %>%
  slice_max(n = top_n_regions_flows, order_by = live_weight_t)

# Create tables
## Top exporter countries
top_exporter_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = exporter_iso3c), by = "rank") %>%
  left_join(top_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = exporter_iso3c), by = "rank")

kable(top_exporter_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_countries, "exporting countries for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top exporter regions
top_exporter_region_table <- data.frame(rank = 1:top_n_regions) %>%
  left_join(top_region_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_total_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_capture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_capture_2019 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_1996 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_1996 = exporter_region), by = "rank") %>%
  left_join(top_region_aquaculture_exporters_2019 %>% 
              mutate(rank = 1:top_n_regions) %>%
              select(rank, top_aquaculture_2019 = exporter_region), by = "rank")

kable(top_exporter_region_table, 
      col.names = c("Rank", "All 1996", "All 2019", 
                    "Capture 1996", "Capture 2019", 
                    "Aqua 1996", "Aqua 2019"),
      caption = paste("Top", top_n_regions, "exporting regions for all blue foods, capture and aquaculture in 1996 versus 2019", sep = " ")) %>%
  kable_styling()

## Top importer countries
top_importer_table <- data.frame(rank = 1:top_n_countries) %>%
  left_join(top_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_total_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_capture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_capture_2019 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_1996 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_1996 = importer_iso3c), by = "rank") %>%
  left_join(top_aquaculture_importers_2019 %>% 
              mutate(rank = 1:top_n_countries) %>%
              select(rank, top_aquaculture_2019 = importer_iso3c), by = "rank")

>>>>

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load packages
library(tidyverse)
library(countrycode)
library(kableExtra)
library(ggpubr)
library(tidytext)
```

```{r load_data}
# Directory
datadir <- "outputs"

global_export_by_source <- read.csv(file.path(datadir, "global_export_by_source.csv"))
all_degree_summary <- read.csv(file.path(datadir, "all_degree_summary.csv"))
habitat_method_degree_summary <- read.csv(file.path(datadir, "habitat_method_degree_summary.csv"))
```

@@

```{r trends_summary}
# Generate stats for text
total_1996 <- round(sum(global_export_by_source %>% filter(year == 1996) %>% pull(live_weight_t))/1000000, 1)
total_2019 <- round(sum(global_export_by_source %>% filter(year == 2019) %>% pull(live_weight_t))/1000000, 1)

@@

# Total trade by environment and production method
# Total trade for marine capture
total_marine_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_capture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland capture")) %>% pull(live_weight_t))/1000000, 1)
# Total trade for marine aquaculture
total_marine_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_marine_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "marine aquaculture")) %>% pull(live_weight_t))/1000000, 1)
# Total trade for inland aquaculture
total_inland_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_inland_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "inland aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for all aquaculture
total_aquaculture_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)
total_aquaculture_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, str_detect(habitat_method, "aquaculture")) %>% pull(live_weight_t))/1000000, 1)

# Total trade for FMFO
total_fmfo_1996 <- round(sum(global_export_by_source %>% filter(year == 1996, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
total_fmfo_2019 <- round(sum(global_export_by_source %>% filter(year == 2019, food_or_fmfo == "fishmeal") %>% pull(live_weight_t))/1000000, 1)
```

```{r percent_export_by_source_trends}
percent_export_by_source <- read.csv(file.path(datadir, "domestic_export_percent_production_by_habitat_method.csv"))
percent_export_total <- read.csv(file.path(datadir, "total_export_percent_prod.csv"))
percent_export_total_dom_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_dom_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "domestic exports") %>% 
  pull(percent_export)
percent_export_total_1996 <- percent_export_total %>% 
  filter(year == 1996, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_export_total_2019 <- percent_export_total %>% 
  filter(year == 2019, calc_type == "all export sources") %>% 
  pull(percent_export)
percent_marine_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine capture") %>%
  pull(percent_export)
percent_marine_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_marine_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "marine aquaculture") %>%
  pull(percent_export)
percent_inland_capture_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_capture_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland capture") %>%
  pull(percent_export)
percent_inland_aqua_exported_1996 <- percent_export_by_source %>% 
  filter(year == 1996, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
percent_inland_aqua_exported_2019 <- percent_export_by_source %>% 
  filter(year == 2019, habitat_method == "inland aquaculture") %>%
  pull(percent_export)
```

```{r trends_concentration}
# Export / Imports concentration
cumulative_percent_threshold <- 75
export_concentration_n_countries <- read.csv(file.path(datadir, "export_concentration_n_countries.csv"))

import_concentration_n_countries <- read.csv(file.path(datadir, "import_concentration_n_countries.csv"))
```

```{r degree_summary_stats}
all_degree_1996 <- all_degree_summary %>%
  filter(year == 1996) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)
all_degree_2019 <- all_degree_summary %>%
  filter(year == 2019) %>% 
  summarise(all_out_degree = mean(all_out_degree, na.rm = TRUE)) %>%
  pull(all_out_degree)
```

@@

![Figure 1: Increases in global export of blue foods. a) Exports of global marine and freshwater aquaculture and fishery products (t live weight equivalent) from 1996-2019. b) Percent of global marine and freshwater aquaculture and fishery production exported, excluding re-exports. c) Out degree by production method and environment.](outputs/fig1.png)

@@

@@

```{r}
bilateral_flow_change <- read.csv(file.path(datadir, "bilateral_flow_change.csv"))

percent_flow_increases <- 100*(nrow(bilateral_flow_change %>% filter(change > 0)))/nrow(bilateral_flow_change)
```

Since 1996, nearly all regional trade pairs increased, other than South America to Europe and Asia and within North America. The largest average annual growth increases occurred for trade within Asia, within Europe and between Europe and Asia. At the country level, `r round(percent_flow_increases, 1)`% of trade pairs increased, with the most increases for marine/inland aquaculture/capture. The declining exports from South America mirror declining marine catch over this period. [Suggestions for discussion points and context related to this?].

![Figure 2: Regional trade flows by trade partner and production source. a) Total imports to each region colored by trade partner. b) Bilateral flows colored by production method and habitat source. c) Total exports from each region colored by trade partner. Values represent million tonnes in live weight equivalents.](outputs/fig2.png)

# National blue food supplies
```{r}
supply <- read.csv(file.path(datadir, "supply.csv"))

global_supply <- read.csv(file.path(datadir, "global_supply.csv"))
